### 2020-06-30 Manuscript/trajectories_manuscript_appendix.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
We fitted a logistic model using the proportion of infants that acquired
each item as output, and age-bin (Defined as we did with our data) as
input, estimating the three parameters afore mentioned (asymptote,
mid-point, and steepness). We included data from 13119 items from 20
languages\footnote{Chinese (Cantonese and Mandarin), Croatian, Czech,
  Danish, English (Australian, British, and American), French
  (European), German, Hebrew, Italian, Korean, Latvian, Portuguese,
  Russian, Slovak, Spanish (European and Mexican), and Turkish. All the
  other available languages were excluded, as no (or insufficient) data
  were available.}, from monolingual infants aged 16 to 30 months of
age.

\textbf{Figure caption}: ``Logarithmic model fitted on Wordbank data,
compared to a linear model. The logarithmic model fits the data slightly
better. The estimated values of the three parameters involved in the
logarithmic model (asymptote, mid-point and steepness) are indicated.''
\end{appendix}