### 2020-07-07 Code/Rmd/model.Rmd
Add SI about the model

---
title: "Model and priors"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Our model specifies a logistic curve that fits our outcome measure. Out outcome measure is the porportion of participants that are reported to understand each word. This logistic curve is characterised by three parameters: *asymptote* ($asym$, the value at which the outcome *pleateaus*), *steepness*, ($steep$, how fast the curve grows), and *mid-point* ($mid-point$, the age at which the curve reaches its maximum steepness). We identify the *mid-point* of a word as its age of acquisition, and thus us the parameter of interest. Out hypothesis is that the language a word belongs to (non-dominant vs. dominant), together with its cognate status (non-cognate vs. cognate) and frequency, affects the position of its mid-point. We assume that the mid-point is generated from a zero-one beta regression model that includes the aforementioned predictors. A beta regression model helps us model data that include many values at boundary, as ours are. We also included random intercepts for each *translation equivalent* ($TE$). We set informed priors for *asymptotes*, *steepnesses*, and *mid-points* according to Wordbank data, and weak priors the the coefficients of the beta-regression model.

$$
\begin{aligned}proportion_{ij} \sim \frac{asym}{1 + e^{(mid_i-age) \times steep}} \\ asym \sim Normal(0.79, 0.1) \\ steep \sim Normal(1.76, 0.8) \\ mid \sim Beta(\mu_i, \phi) \\ \mu_i = \beta_0 + TE_{0i} + \beta_1 \times dominance_i + \beta_2 \times cognateness_i \\ \beta_0 \sim Normal(4.50, 1) \\ \beta_1 \sim Normal(0, 10) \\ \beta_2 \sim Normal(0, 10) \\ \phi \sim Normal(1.5, 1)
\end{aligned}
$$