### 2021-01-13 Rmd/04_appendix_2.Rmd
Merge manuscript sections

---
title: 'Appendix 2: Model'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```