### 2021-07-12 Rmd/02_methods.Rmd
I hope I don't overwrite remote repo

---
title: "Methods"
author: "Gonzalo García-Castro"
date: "6/22/2021"
output:
  word_document:
    reference_docx: trajectories_template.docx
  pdf_document:
      latex_engine: xelatex
bibliography: trajectories.bib
csl: apa6.csl
self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Methods

## Participants

## Vocabulary checklist

## Data analysis

$$

$$