### 2021-09-15 Rmd/report.Rmd
Fix figures

---
title: "Trajectories"
output:
  github_document:
    pandoc_args: --webtex
bibliography: "trajectories.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = TRUE,
  echo = TRUE,
  warning = TRUE,
  dev.args = list(png = list(type = "cairo")),
  include = TRUE,
  out.width = "100%",
  results = "asis",
  dpi = 300
)
options(
  knitr.kable.NA = "-",
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
  ggplot2.continuous.fill = ggplot2::scale_color_gradient,
  ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# load packages
library(targets)
library(dplyr)
library(tidyr)
library(stringr)
library(gt)
library(lubridate)
library(wesanderson)
library(brms)
library(htmltools)
library(tidybayes)
library(scales)
library(patchwork)
library(tibble)
library(janitor)
library(here)

# set params
tar_load_globals()
theme_set(theme_custom())

# import data
tar_load(items)
tar_load(participants)
tar_load(vocabulary)
tar_load(responses)
tar_load(fits)

```

# Questionnaires

# Items {.tabset .tab-pills}

```{r items_list, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=4, out.width="100%"}
items %>% 
  pivot_wider(
    id_cols = c("te", "class", "cognate"),
    names_from = "language",
    values_from = c("label", "frequency", "ipa")
  ) %>% 
  clean_names() %>% 
  arrange(te) %>% 
  gt() %>% 
  tab_spanner(label = md("**Catalan**"), columns = contains("_catalan")) %>% 
  tab_spanner(label = md("**Spanish**"), columns = contains("_spanish")) %>% 
  cols_label(
    te = "TE ID",
    class = "Class",
    cognate = "Cognateness",
    label_catalan = "Label",
    label_spanish = "Label",
    ipa_catalan = "Phonology",
    ipa_spanish = "Phonology",
    frequency_catalan = "Frequency",
    frequency_spanish = "Frequency"
  ) %>% 
  fmt_number(columns = starts_with("frequency")) %>% 
  as_raw_html() %>% 
  div(style = "height:600px; overflow-x: scroll; overflow-y: scroll", .)
```

<br>

## Lexical frequency

```{r items_frequency, message=FALSE, warning=FALSE, paged.print=FALSE}

by_cognate <- items %>% 
  drop_na(frequency) %>% 
  ggplot(aes(frequency, cognate, fill = cognate, colour = cognate)) +
  geom_jitter(height = 0.05, alpha = 0.1, shape = 1, stroke = 1) +
  stat_slab(position = position_nudge(y = 0.1), alpha = 0.5) +
  geom_boxplot(width = 0.05, colour = "black", fill = "white",
               position = position_nudge(y = 0.1), outlier.colour = NA) +
  labs(
    y = "Cognateness", 
    x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
    colour = "cognateness", 
    fill = "Cognateness"
  )

by_language <- items %>% 
  drop_na(frequency) %>% 
  ggplot(aes(frequency, language, fill = cognate, colour = cognate)) +
  geom_jitter(height = 0.05, alpha = 0.1, shape = 1, stroke = 1) +
  stat_slab(position = position_nudge(y = 0.1), alpha = 0.5) +
  geom_boxplot(width = 0.05, colour = "black", fill = "white",
               position = position_nudge(y = 0.1), outlier.colour = NA) +
  labs(y = "Cognateness", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness")

by_class <- items %>% 
  drop_na(frequency, class) %>% 
  ggplot(aes(frequency, class, fill = cognate, colour = cognate)) +
  geom_jitter(height = 0.05, alpha = 0.1, shape = 1, stroke = 1) +
  stat_slab(position = position_nudge(y = 0.1), alpha = 0.5) +
  geom_boxplot(width = 0.05, colour = "black", fill = "white",
               position = position_nudge(y = 0.1), outlier.colour = NA) +
  labs(y = "Cognateness", x = "Lexical frequency", colour = "cognateness", fill = "Cognateness")

(by_cognate / by_language) + by_class +
  plot_layout(2) &
  theme_custom() &
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    panel.grid.major.y = element_blank()
  )

```

## Cognateness

```{r items_cognate, message=FALSE, warning=FALSE, paged.print=FALSE}
by_cognate <- items %>% 
  distinct(te, class, cognate) %>% 
  ggplot(aes(x = cognate, fill = class)) +
  geom_bar() +
  coord_flip() +
  labs(x = "Cognateness", y = "Number of items", fill = "Class")

by_class <- items %>% 
  distinct(te, class, cognate) %>% 
  ggplot(aes(x = class, fill = cognate)) +
  geom_bar(position = position_fill()) +
  labs(x = "Cognateness", y = "Proportion of items (%)", fill = "Cognateness") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent)

by_cognate / by_class +
  plot_layout() &
  theme_custom() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right",
    legend.title = element_blank()
  )

```

## {-}

# Participants {.tabset .tab-pills}

```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
  group_by(time_stamp, lp, dominant_language) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  group_by(lp, dominant_language) %>% 
  mutate(n = cumsum(n)) %>% 
  ggplot(aes(time_stamp, n, colour = interaction(lp, dominant_language, sep = " - "), fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_line(size = 1) +
  labs(x = "Date", y = "Number of responses", colour = "LP - Dominant language") +
  scale_y_continuous(breaks = seq(0, 250, 50)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
  )

```

## Age

```{r participants_age, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>% 
  mutate(age = round(age)) %>% 
  count(lp, dominant_language, age) %>% 
  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_col(position = position_fill(), colour = "white") +
  guides(colour = guide_legend(ncol = 2)) +
  coord_flip() +
  labs(y = "Proportion of the sample", x = "Age (months)", colour = "LP - Dominant language") +
  scale_y_continuous(labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
  ) +

  participants %>% 
  mutate(age = round(age)) %>% 
  count(lp, dominant_language, age) %>% 
  ggplot(aes(as.factor(age), n, fill = interaction(lp, dominant_language, sep = " - "))) +
  geom_col(colour = "white") +
  labs(y = "Number of participants", x = "Age (months)") +
  guides(colour = guide_legend(ncol = 2)) +
  theme_custom() +
  coord_flip() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank()
  ) +

  plot_layout(nrow = 1, guides = "collect") &
  theme(
    legend.position = "top", 
    axis.text.y = element_text(size = 9),
    panel.grid.major = element_blank()
  )

```

## Language profile

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
a <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  ) %>% 
  as_raw_html()

```

## {-}

# Data analysis {.tabset .tab-pills}

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age` by `item`, and their correlation parameter. We then extended this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`. The models were implemented in `brms` as:

* `understands/produces ~ 1 + age + frequency + (1 + age | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age + frequency + doe + (1 + age + doe | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age + frequency + doe*cognate + (1 + age + doe | item) + (1 + frequency + cognate| id)`

## Model equation

$$
\begin{aligned}
\color{white}
\log(\frac{p}{n - q}) &= (\beta_{0} + \beta_{0p} + \beta_{0i}) + ... \\
& (\beta_{1} + \beta_{1p})\times Cognate_{pi} + ... \\
& (\beta_{2} + \beta_{2i}) \times Exposure_{pi} + ... \\
& (\beta_{3} + \beta_{3i} + \beta_{3p}) \times (Cognate_{pi} \times Exposure_{pi}) + ... \\
& (\beta_{4} + \beta_{4i}) \times Age_{pi} + ... \\
& (\beta_{4} + \beta_{4p}) \times Frequency_{pi} + ... \\
& \varepsilon_{pi}
\end{aligned}
$$

## R code (`brms`)

```
understands ~
1 + age + frequency + lp*cognate +
(1 + age + lp | te) +
(1 + frequency + cognate | id),
family =  bernoulli("logit")
```

## Stan code

Stan code generated by `brms::stancode`:

```{stan stan_model, output.var="model"}
// generated with brms 2.15.0
functions {
/* compute correlated group-level effects
* Args: 
*   z: matrix of unscaled group-level effects
*   SD: vector of standard deviation parameters
*   L: cholesky factor correlation matrix
* Returns: 
*   matrix of scaled group-level effects
*/ 
matrix scale_r_cor(matrix z, vector SD, matrix L) {
// r is stored in another dimension order than z
return transpose(diag_pre_multiply(SD, L) * z);
}
}
data {
int<lower=1> N;  // total number of observations
int Y[N];  // response variable
int<lower=1> K;  // number of population-level effects
matrix[N, K] X;  // population-level design matrix
// data for group-level effects of ID 1
int<lower=1> N_1;  // number of grouping levels
int<lower=1> M_1;  // number of coefficients per level
int<lower=1> J_1[N];  // grouping indicator per observation
// group-level predictor values
vector[N] Z_1_1;
vector[N] Z_1_2;
vector[N] Z_1_3;
int<lower=1> NC_1;  // number of group-level correlations
// data for group-level effects of ID 2
int<lower=1> N_2;  // number of grouping levels
int<lower=1> M_2;  // number of coefficients per level
int<lower=1> J_2[N];  // grouping indicator per observation
// group-level predictor values
vector[N] Z_2_1;
vector[N] Z_2_2;
vector[N] Z_2_3;
int<lower=1> NC_2;  // number of group-level correlations
int prior_only;  // should the likelihood be ignored?
}
transformed data {
int Kc = K - 1;
matrix[N, Kc] Xc;  // centered version of X without an intercept
vector[Kc] means_X;  // column means of X before centering
for (i in 2:K) {
means_X[i - 1] = mean(X[, i]);
Xc[, i - 1] = X[, i] - means_X[i - 1];
}
}
parameters {
vector[Kc] b;  // population-level effects
real Intercept;  // temporary intercept for centered predictors
matrix[M_1, N_1] z_1;  // standardized group-level effects
cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
vector<lower=0>[M_2] sd_2;  // group-level standard deviations
matrix[M_2, N_2] z_2;  // standardized group-level effects
cholesky_factor_corr[M_2] L_2;  // cholesky factor of correlation matrix
}
transformed parameters {
vector<lower=0>[M_1] sd_1;  // group-level standard deviations
matrix[N_1, M_1] r_1;  // actual group-level effects
// using vectors speeds up indexing in loops
vector[N_1] r_1_1;
vector[N_1] r_1_2;
vector[N_1] r_1_3;
matrix[N_2, M_2] r_2;  // actual group-level effects
// using vectors speeds up indexing in loops
vector[N_2] r_2_1;
vector[N_2] r_2_2;
vector[N_2] r_2_3;
sd_1 = rep_vector(1, rows(sd_1));
// compute actual group-level effects
r_1 = scale_r_cor(z_1, sd_1, L_1);
r_1_1 = r_1[, 1];
r_1_2 = r_1[, 2];
r_1_3 = r_1[, 3];
// compute actual group-level effects
r_2 = scale_r_cor(z_2, sd_2, L_2);
r_2_1 = r_2[, 1];
r_2_2 = r_2[, 2];
r_2_3 = r_2[, 3];
}
model {
// likelihood including constants
if (!prior_only) {
// initialize linear predictor term
vector[N] mu = Intercept + rep_vector(0.0, N);
for (n in 1:N) {
// add more terms to the linear predictor
mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_2_1[J_2[n]] * Z_2_1[n] + r_2_2[J_2[n]] * Z_2_2[n] + r_2_3[J_2[n]] * Z_2_3[n];
}
target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
}
// priors including constants
target += normal_lpdf(b | 0, 3);
target += normal_lpdf(Intercept | 0, 3);
target += std_normal_lpdf(to_vector(z_1));
target += lkj_corr_cholesky_lpdf(L_1 | 5);
target += normal_lpdf(sd_2 | 0, 0.5)
- 3 * normal_lccdf(0 | 0, 0.5);
target += std_normal_lpdf(to_vector(z_2));
target += lkj_corr_cholesky_lpdf(L_2 | 5);
}
generated quantities {
// actual population-level intercept
real b_Intercept = Intercept - dot_product(means_X, b);
// compute group-level correlations
corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
vector<lower=-1,upper=1>[NC_1] cor_1;
// compute group-level correlations
corr_matrix[M_2] Cor_2 = multiply_lower_tri_self_transpose(L_2);
vector<lower=-1,upper=1>[NC_2] cor_2;
// extract upper diagonal of correlation matrix
for (k in 1:M_1) {
for (j in 1:(k - 1)) {
cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
}
}
// extract upper diagonal of correlation matrix
for (k in 1:M_2) {
for (j in 1:(k - 1)) {
cor_2[choose(k - 1, 2) + j] = Cor_2[j, k];
}
}
}
```

## {-}

# Results {.tabset .tab-pills}

## Vocabulary {.tabset .tab-pills}

### Vocabulary size {.tabset .tab-pills}

#### Comprehension

```{r vocabulary_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}

d <- vocabulary$wide %>% 
  filter(type=="Understands", modality != "Te")

ggplot(d, aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  ggplot(d, aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

  plot_layout(guides = "collect") &
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

```

#### Production

```{r vocabulary_production, message=FALSE, warning=FALSE, paged.print=FALSE}

vocabulary$wide %>% 
  filter(type=="Produces", modality != "Te") %>% 
  ggplot(aes(age, prop, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (%)", colour = "LP", fill = "LP") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +

  vocabulary$wide %>% 
  filter(type=="Produces", modality != "Te") %>% 
  ggplot(aes(age, count, colour = lp, fill = lp)) +
  facet_wrap(~modality) +
  geom_point(size = 0.5, alpha = 0.21) +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = FALSE) +
  labs(x = "Age (months)", y = "Vocabulary size (words)", colour = "LP", fill = "LP") +

  plot_layout(guides = "collect") &
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

```

#### {-}

### Short vs. Long {.tabset .tab-pills}

#### Comprehension

```{r vocabulary_long_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

#### Production

```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="produces") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

#### {-}

### {-}

## Raw data

```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  group_by(te, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(prop, reorder(te, -prop), colour = cognate)) +
  facet_wrap(vars(type)) +
  geom_vline(xintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID", colour = "Type") +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(age = as.factor(round(age))) %>% 
  group_by(te, age, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(age, prop, colour = cognate, fill = cognate, shape = cognate)) +
  facet_wrap(vars(type)) +
  geom_hline(yintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(aes(group = interaction(age, cognate)), alpha = 0.5, size = 0.5) +
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID",
       colour = "Cognate", shape = "Cognate", fill = "Cognate") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank()
  )
```

## Prior-predictive checks

```{r ppc, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate = unique(items$cognate),
  frequency_center = c(-1, 0, 1)
)

m <- add_fitted_draws(nd, fits$fit_2, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

```

## Model selection

We compared the performance of these models using Bayesian leave-one-out cross-validation (LOO) using the `loo` and `loo_compare` functions of the `brms` R package (dependent of the `LOO` R package). LOO consists in computing the average likelihood of each observation after estimating the model's parameters leave that same observation out of the data set. Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.

## Fixed effects

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe", "cognate1", "doe:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

```

<br>

```{r fix_effets_plot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI. The intercept has been transformed using the inverse logit to get the average probability of correct response. The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}

post <- gather_draws(fits$fit_2, `b_.*`, regex = TRUE) %>% 
  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))

ggplot(post, aes(.value)) +
  facet_wrap(~.variable, scales = "free_x") +
  geom_vline(xintercept = 0, size = 1, colour = "grey") +
  stat_histinterval(fill = "lightblue") +
  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
  theme_custom() +
  theme(
    legend.position = "none",
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0),
    # axis.text = element_text(size = 7),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )
```

## Random effects {.tabset .tab-pils}

### Participant

```{r random_participant, message=FALSE, warning=FALSE, paged.print=FALSE}

re_id <- ranef(fits$fit_2)$id[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("id") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
  left_join(distinct(fits$fit_2$data, id, lp))

ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[1]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Participant average probability of comprehension",
       y = "Participant") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_custom() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

```

### Item

```{r re_item, message=FALSE, warning=FALSE, paged.print=FALSE}

re_te <- ranef(fits$fit_2)$te[,,1] %>% 
  as.data.frame() %>% 
  rownames_to_column("te") %>% 
  as_tibble() %>% 
  mutate_if(is.numeric, function(x) inv_logit_scaled(x + fixef(fits$fit_2)[1])) %>% 
  mutate(te = as.numeric(te)) %>% 
  left_join(distinct(fits$fit_2$data, te, lp))

ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[2]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Item average probability of comprehension",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_custom() +
  theme(
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  )
```

### {-}

## Marginal means

```{r emmeans, message=FALSE, warning=FALSE, paged.print=FALSE}

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate =unique(items$cognate),
  frequency_center = c(-1,  0, 1)
)

m <- add_fitted_draws(nd, fits$fit_2, re_formula = NA, n = 25) %>% 
  mutate(frequency_center = as.factor(paste0(frequency_center, " SD")))

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_wrap(lp~frequency_center) +
  geom_line(aes(group = interaction(.draw, frequency_center, cognate)), alpha = 0.5) +

  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_custom() +
  theme(
    legend.position = "right",
    legend.title = element_blank()
  )

```

### Area under the curve (AUC)

### Traceplots

```{r diagnostics_traceplots, message=FALSE, warning=FALSE, paged.print=FALSE}

gather_draws(fits$fit_2, `b_.*`, `sd_te__.*`, `cor_te__.*`, regex = TRUE) %>% 
  mutate(.chain = paste("Chain ", .chain)) %>% 
  ggplot(aes(.iteration, .value, colour = .chain)) +
  facet_wrap(~.variable, scales = "free_y") +
  annotate(geom = "rect", colour = NA, xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.25, colour = "grey") +
  geom_line(alpha = 0.5) +
  labs(x = "Iteration", y = "Sample value", colour = "Chain") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid = element_blank()
  )

```

## {-}