### 2021-09-24 Rmd/03_results.Rmd

---
title: "Results"
output:
  word_document:
    reference_docx: trajectories_template.docx
  pdf_document:
      latex_engine: xelatex
bibliography: trajectories.bib
csl: apa6.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r results, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
tar_load(fits_comp)
tar_load(fits_prod)
tar_load(loo_comp)
tar_load(loo_prod)
```

## Comprehension

The extended model, including the three-way interaction `dominance1:bilingualism;cognate1` fitted comprehension data the best: it showed the smallest absolute expected log predictive density. This estimate is several times larger than its standard error indicating high predictive accuracy. Table 2 summarises the LOO scores of the fitted models.

We report the mean of each coefficient's posterior distribution along with standard errors and 95% CrI in Table 3. For interpretability, we report de derivative of the logistic function^[The derivative is calculated as $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$. This value corresponds to the slope of the logistic curve at the midpoint, where it is steepest.] for each coefficient, which indicates the maximum difference in probability corresponding to one unit different in the input variable [@gelman_regression_2020].

The intercept (`r printnum(inv_logit_scaled(coef_comp["Intercept", "Estimate"])*100)`%) indicates probability of participants understanding TEs at mean age (`r printnum(mean(responses$age_scaled))` months), mean frequency (`r printnum(mean(responses$frequency_scaled))` Zipf score), and mean L2 exposure (`r printnum(mean(responses$bilingualism_scaled))`%), across the levels of the `dominance` and `cognate` factors. An increase in one standard deviation increment in age (1 *SD* = `r printnum(sd(responses$age_scaled))` months) is associated with a `r printnum(coef_comp["age", "Estimate"]*100/4)`% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 *SD* = `r printnum(sd(responses$frequency_scaled))` Zipf scores) is associated with a `r printnum(coef_comp["frequency", "Estimate"]*100/4)`% increment in comprehension probability. Items in the dominant language were `r printnum(coef_comp["dominance1", "Estimate"]*100/4)`% more likely to be understood than items in the non-dominant language. A one standard deviation increase in exposure to the non-dominant language (1 *SD* = `printnum(sd(responses$bilingualism_scaled))`%) is associated with a `r printnum(coef_comp["bilingualism", "Estimate"]*100/4)` increment in comprehension probability. Cognates were `r printnum(coef_comp["cognate1", "Estimate"]*100/4)`% more like to be understood than non-cognates. The 95% credible interval of the three-way interaction (`dominance:bilingualism:cognate`) excluded 0, indicating that that the cognate status of the items changed the effect of dominance differently depending on the amount of exposure to the second language. Figure 1 shows the estimated marginal mean posterior probability of comprehension and production across ages, and the different levels of `dominance`, `bilingualism`, and `cognate`.

Children exposed to their second language more than -1 SD away from the mean (`r printnum(mean(responses$bilingualism_scaled)-sd(responses$bilingualism_scaled))`%) were more likely to understand cognates than non-cognates in their non-dominance language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[6]*100)`%, 95% CrI = [`r printnum(deltas$.lower[6]*100)`-`r printnum(deltas$.upper[6]*100)`]), but not in their dominant language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[2]*100)`%, 95% CrI = [`r printnum(deltas$.lower[2]*100)`-`r printnum(deltas$.upper[2]*100)`]). This difference was smaller for participants with mean exposure to the non-dominant language (`r printnum(mean(responses$bilingualism_scaled))`%): in the non-dominant language cognates were slight more likely to be understood than non-cognates ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[12]*100)`%, 95% CrI = [`r printnum(deltas$.lower[12]*100)`-`r printnum(deltas$.upper[12]*100)`]), but not in the dominant language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[8]*100)`%, 95% CrI = [`r printnum(deltas$.lower[8]*100)`-`r printnum(deltas$.upper[8]*100)`]). Finally this effect was negligible for participants exposed to their second language more than -1 SD away from the mean (`r printnum(mean(responses$bilingualism_scaled)+sd(responses$bilingualism_scaled))`%). Cognates and non-cognates were equally likely to be understood in both the non-dominant ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[18]*100)`%, 95% CrI = [`r printnum(deltas$.lower[18]*100)`-`r printnum(deltas$.upper[18]*100)`]), and the dominant language ($M_{Cognate}-M_{Non-cognate}$ = `r printnum(deltas$delta[14]*100)`%, 95% CrI = [`r printnum(deltas$.lower[14]*100)`-`r printnum(deltas$.upper[14]*100)`]).

The difference in probability of comprehension between the dominant and the non-dominant languages differed between cognates and non-cognates and different levels of exposure to the non-dominant language. For children exposed. For children exposed to their non-dominant language less than 1 SD from the mean, difference in probability between the dominant and the non-dominant language was smaller for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[4]*100)`%, 95% CrI = [`r printnum(deltas$.lower[4]*100)`-`r printnum(deltas$.upper[4]*100)`]) than for non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[1]*100)`%, 95% CrI = [`r printnum(deltas$.lower[1]*100)`-`r printnum(deltas$.upper[1]*100)`]). For children with mean exposure to the non-dominant language, this effect was smaller: the difference between the comprehension probability of items in L1 and L2 was slightly smaller for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[10]*100)`%, 95% CrI = [`r printnum(deltas$.lower[10]*100)`-`r printnum(deltas$.upper[10]*100)`]) than for non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[7]*100)`%, 95% CrI = [`r printnum(deltas$.lower[7]*100)`-`r printnum(deltas$.upper[7]*100)`]). Finally this effect was negligible for children exposed more than 1 SD than the mean to their non-dominant language: the difference in comprehension probability between items in the dominant and the non-dominant language was roughly similar for cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[16]*100)`%, 95% CrI = [`r printnum(deltas$.lower[16]*100)`-`r printnum(deltas$.upper[16]*100)`]) and non-cognates ($M_{Dominant}-M_{Non-dominant}$ = `r printnum(deltas$delta[13]*100)`%, 95% CrI = [`r printnum(deltas$.lower[13]*100)`-`r printnum(deltas$.upper[13]*100)`]).