### 2022-11-17 docs/childes-freqs.qmd
Update lab notes

---
title: "Cross/language CHILDES frequencies"
format: html
code-fold: true
---

```{r}
#| label: setup
#| echo: true
library(childesr)
library(dplyr)
library(ggplot2)
library(purrr)
library(patchwork)
library(tidyr)
library(janitor)
library(ggdist)
library(gt)
library(arrow)
library(readr)
```

```{r}
#| label: wordbank
#| echo: true
# function to retrieve frequencies
wordbank <- map_df(
    c(spa = "data/wordbank_spa.csv", eng = "data/wordbank_eng.csv"),
    read_csv,
    .id = "language"
    ) %>% 
    clean_names() %>% 
    select(language, age, definition, uni_lemma, ipa, length, freq)

wordbank %>%
    drop_na() %>% 
    distinct(uni_lemma, definition, language, freq) %>% 
    arrange(uni_lemma) %>% 
    pivot_wider(
        id_cols = uni_lemma,
        names_from = language,
        values_from = c(freq, definition),
        values_fn = first
        ) %>% 
    drop_na() %>% 
    ggplot(aes(freq_eng, freq_spa)) +
    geom_point()
```

```{r}
#| label: retrieve-freqs
#| echo: true
# function to retrieve frequencies

get_childes_data <- function(word, collection, ...){

    roles <- c("Mother", "Father", "Investigator", "Sibling", "Sister", "Grandmother", "Adult", "Friend", "Brother", "Visitor", "Relative", "Grandfather", "Teacher", "Student")

    counts <- get_types(type = word, collection = collection, role = roles, ...)

    speaker_ids <- distinct(counts, collection_id, corpus_id, transcript_id, speaker_id)

    speakers <- speaker_ids %>%
        left_join(get_speaker_statistics(collection = collection)) %>%
        select(collection_id, corpus_id, transcript_id, speaker_id, num_tokens)

    count_estimates <- counts %>%
        left_join(speakers) %>%
        mutate(
            id = as.character(id),
            age_months = target_child_age,
            age_bin = as.integer(floor(age_months / 6) * 6),
            token = tolower(gloss)
        ) %>%
        group_by(age_bin, token, target_child_id, transcript_id) %>%
        summarise(
            transcript_count = sum(count),
            transcript_num_tokens = sum(num_tokens),
            .groups = "drop"
        ) %>%
        group_by(age_bin, token) %>%
        summarise(
            freq_count = mean(transcript_count), 
            total_count = mean(transcript_num_tokens),
            n_children = length(unique(target_child_id)),
            .groups = "drop"
        ) %>%
        mutate(
            freq_million = freq_count / total_count * 1e6,
            freq_zipf = log10(freq_million)+3
        )

    return(count_estimates)

}

# params
tokens <- list(
    spa = c("casa", "puerta", "mesa", "gato", "caja", "lampara"),
    cat = c("casa", "porta", "taula", "gat", "caixa", "llum"),
    eng = c("house", "door", "table", "cat", "box", "lamp")
)

freqs <- map2_df(
    collections, 
    corpora, 
    ~get_frequencies(collection = .x, corpus = .y, age = c(10, 36)), 
    .id = "language"
)
```

# How many children, transcriptions, and tokens are available in each language?

```{r}
#| label: num-children
#| echo: true
collections <- list(
    spa = "Spanish",
    cat = "Romance",
    eng = c("Eng-NA", "Eng-UK")
)

corpora <- list(
    spa = NULL,
    cat = c("EsteveGibert/Prieto", "GRERLI", "Jordina", "Julia", "Mireia/Eva/Pascual", "Serra/Solé"),
    eng = NULL
)

counts <- map2_df(
    collections, 
    corpora, 
    ~get_types(collection = .x, role = roles, corpus = .y),
    .id = "language"
)

counts_to_transpose <- counts %>% 
    drop_na(target_child_id) %>% 
    group_by(language) %>% 
    summarise(
        n_children = length(unique(target_child_id)),
        n_transcripts = length(unique(transcript_id)),
        n_distinct_words = length(unique(gloss)),
        n_word_instances = sum(count),
        median_word_instances = median(count),
        max_word_instances = max(count)
    )

as_tibble(cbind(nms = names(counts_to_transpose), t(counts_to_transpose))) %>% 
    set_names(.[1,]) %>% 
    rename(variable = language) %>% 
    slice(-1) %>% 
    mutate(across(cat:spa, ~as.numeric(str_trim(.)))) %>% 
    gt() %>% 
    fmt_integer(2:4)

```

ggplot(freqs, aes(age_bin, freq_zipf)) +
    geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) +
    scale_y_continuous(limits  = c(1, 7))

tes <- tibble(
    language = c("cat", "spa", "eng"),
    house = c("casa", "casa", "house")
)

freqs %>% 
    filter(
        (language=="cat" & token=="casa") |
            (language=="spa" & token=="casa") |
            (language=="eng" & token=="house")
    ) %>% 
    ggplot(aes(age_bin, freq_zipf, colour = language)) +
    geom_point(size = 4, shape = 1, stroke = 2) +
    scale_y_continuous(limits = c(1, 7))

group_by(age_bin, language) %>% 
    summarise(total_count = sum(total_count)) %>% 
    ggplot(aes(age_bin, total_count, fill = language)) +
    geom_col(position = position_dodge(), width = 3) +
    geom_text(
        aes(label = round(total_count), y = total_count + 100),
        position = position_dodge(width = 1),
        angle = 90,
        nudge_y = 100
    )

freqs %>% 
    ggplot() +
    aes(
        x = age_bin,
        y = freq_zipf,
        colour = language
    ) +
    facet_wrap(~token) +
    stat_summary(fun.data = "mean_se", geom = "pointrange") +
    facet_wrap(~token, scales = "free_y") +
    scale_x_continuous(
        breaks = seq(10, 36, 4),
        name = "Age (months)"
    ) +
    scale_y_continuous(name = "Frequency (per million words)")

```