### 2022-11-17 index.qmd
Update lab notes

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
authors:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    corresponding: true
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias fargas 25-27
        city: Barcelona
        state: Spain
        postal-code: 08005
  - name: Daniela S. Ávila-Varela
    orcid: 0000-0002-3518-8117
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations:
      - name: Universidad Atónoma de Madrid
        department: Departamento de Psicología
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
format:
    html:
        toc: true
        link-external-newwindow: true
        number-sections: true
        number-offset: 0
        code-fold: true
        code-overflow: scroll
        code-line-numbers: true
        code-copy: hover
        fig-dpi: 1000
        reference-location: margin
        execute-dir: "C:/Users/U155880/Documents/trajectories/"
    pdf:
        toc: true
        number-sections: true
        colorlinks: true
        fig-dpi: 1000
bibliography: "references.bib"
csl: "manuscript/apa6.csl"
---

```{r}
#| label: setup
#| include: false
tar_load_globals() # load packages and options
tar_load_everything() # load targets as R objects
theme_set(theme_custom()) # set custom ggplot theme

# resolve namespace conflicts
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)
```

# Methods

Data and materials are available at OSF (https://osf.io/hy984/), and code is available at GitHub (https://github.com/gongcastro/trajectories).

## Questionnaires

```{r}
#| label: questionnaire
#| echo: true
#| message: false
#| warning: false
pool <- mutate(bvqdev::pool, language = ifelse(str_detect(item, "cat_"), "Catalan", "Spanish"))
n_categories <- distinct(pool, category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% 
    pull(n) %>%
    set_names(c("catalan", "spanish"))
```

The questionnaire was implemented on-line using the formR platform [@arslan2020formr], and was structured in three blocks: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish. The Catalan inventory contained `r n_item_language["catalan"]` items and the Spanish inventory contained `r n_item_language["spanish"]` items. Items in one language were translation equivalents of the items in the other language (e.g., whenever *gos* \[dog\] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping. When there were two acceptable translation equivalents for a given word, we included both in separate items (e.g., Catalan *acabar* \[*to finish*\] and Spanish *acabar* and *terminar*), or merged them into a single items (e.g., Spanish *mono* \[*monkey*\] and Catalan *mono/mico*. We included items from a diverse sample of `r n_categories` semantic/functional categories (see Appendix 1). For the analyses included in this study, we excluded items from the adverbs, auxiliary words, connectives, interjections and games and routines categories, so that only data from content words (nouns, adjectives, and verbs) were used.

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default). Some families filled a long version of the vocabulary checklists (800 translation equivalents; 800 items in Catalan, 800 items in Spanish), while others filled a shorter version (\~400 translation equivalents, \~400 items in Catalan, \~400 items in Spanish). These last families were randomly allocated into one of four different subsets of the complete list of items. These lists were carefully designed so that each contained a representative subsample of the items from the complete list. Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists--were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions. Table 2 in Appendix 1 shows the distribution of items across questionnaire versions. We excluded from the analysis multi-word items (e.g., *barrita de cereales* \[cereal bar\]) and items that included more than one word-form (e.g., *mono / mico*). Table 3 shows the classification of items in cognates and non-cognates and their lexical frequency scores across the four lists of the inventories.

## Items

```{r}
#| label: tbl-items-summary
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(pool, te, item, language, class, category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq_zipf, n_phon, lv), 
        list(
            n = ~sum(!is.na(.)),
            mean = ~mean(., na.rm = TRUE),
            sd = ~sd(., na.rm = TRUE),
            min = ~min(., na.rm = TRUE),
            max = ~max(., na.rm = TRUE)
        )
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_zipf_n) %>% 
    gt(groupname_col = "language") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
    cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>% 
    cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>% 
    cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>% 
    cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>% 
    cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>% 
    cols_merge_range(col_begin = lv_min, col_end = lv_max) %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_zipf_mean = md("Mean (SD)"),
        freq_zipf_min = md("Range"),
        n_phon_mean = md("Mean (SD)"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean (SD)"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    ) %>% 
    summary_rows(
        fns = list("Grand mean" = ~mean(.)),
        columns = ends_with("_mean")
    ) %>% 
    summary_rows(
        fns = list("N" = ~sum(.)),
        columns = n,
        formatter = fmt_integer
    )
```

### Lexical frequency

```{r}
#| label: fig-items-freq
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of lexical frequency across items"
items %>% 
    ggplot() +
    aes(
        x = language,
        y = freq_zipf,
        fill = language,
        colour = language
    ) +
    stat_slab(slab_colour = "white", slab_size = 1, alpha = 0.5, scale = 0.5) +
    stat_dots(side = "left", scale = 0.5, position = position_nudge(x = -0.05)) + 
    stat_pointinterval(
        position = position_dodge(width = 0.1),
        point_size = 2.5,
        colour = "black"
    ) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    theme(
        legend.position = "none",
        legend.title = element_blank()
    )
```

### \# Phonemes

```{r}
#| label: fig-items-nphon
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    count(language, n_phon) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = n,
        fill = language,
        colour = language,
        shape = language
    ) +
    facet_wrap(~language, ncol = 1) +
    geom_col() +
    labs(
        x = "# Phonemes",
        y = "Number of words",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_x_continuous(breaks = 1:15) +
    theme(
        legend.position = "none",
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )
```

### Levenshtein

```{r}
#| label: fig-items-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    ggplot(aes(x = lv)) +
    geom_dots(fill = ggsci::pal_d3()(4)[1], colour = ggsci::pal_d3()(4)[1]) +
    labs(x = "Levenshtein distance", y = "Probability density") +
    scale_x_continuous(labels = percent) +
    theme(
        legend.position = "top"
    )
```

### Frequency by \# Phonemes

```{r}
#| label: fig-items-frequency-phonemes
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq_zipf) %>% 
    count(n_phon, freq_zipf, language) %>% 
    ggplot(aes(n_phon, freq_zipf, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 2, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    coord_equal(ratio = 10) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

### Frequency by Levenshtein

```{r}
#| label: fig-items-frequency-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq_zipf) %>% 
    count(freq_zipf, lv, language) %>% 
    ggplot(aes(freq_zipf, lv, colour = language, fill = language, size = n)) +
    facet_wrap(~language) +
    geom_point(size = 2, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

### \# Phonemes by Levenshtein

```{r}
#| label: fig-items-nphon-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen number of phonemes and phonological similarity"
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq_zipf) %>% 
    count(n_phon, lv, language) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language, size = n)) +
    facet_wrap(~language) +
    geom_point(shape = 1, stroke = 1) +
    geom_smooth(method = "lm", formula = "y ~ x", size = 1) +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

## Participants

```{r}
#| label: fig-participants-time
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Data at which responses were collected"
participants %>% 
    group_by(time_stamp) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 2, colour = pal_d3()(1)[1]) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
    scale_x_date(
        date_breaks = "3 month", 
        date_labels = "%B %Y"
    ) +
    guides(colour = guide_legend(ncol = 2)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 20, hjust = 1),
        axis.title.x = element_blank(),
    )
```

### Age

```{r}
#| label: fig-participants-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

### Language profile

```{r}
#| label: fig-participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
participants %>% 
    mutate(
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
    labs(
        x = "Degree of Exposure (DoE)",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        legend.position = "top",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank()
    )
```

### SES/parental education

```{r}
#| label: tbl-participants-edu
#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
#| echo: true
#| message: false
#| warning: false
participants %>%
    left_join(bvq_data$logs) %>% 
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.2), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(bvq_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    summary_rows(
        columns = Complementary:Primary, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_integer
    ) %>% 
    cols_label(doe_catalan = "DoE Catalan") %>%
    sub_missing(columns = 2:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", Complementary:Primary) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:5)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

## Exposure-weighted lexical frequency

We created a predictor ($Exposure$) to account for the exposure rate of each child to each of the words their parents responded to, weighted by the child's exposure to the language the word belongs to. This measure is the product of the lexical frequency of the word (Zipf score) and the child's degree of exposure to the corresponding language (a proportion):

$$
\textbf{Exposure}_{ij} = \textbf{Frequency}_i \times \textbf{DoE}_{j}
$$

For instance, for a child who is reportedly exposure to Catalan 80% of the time, and to Spanish 20% of the time, the expected exposure to the word *cavall* (horse, in Catalan, with a lexical frequency of `r round(items$freq_zipf[items$item=="cat_cavall"], 2)`) would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.80, 2)`, while that of its translation to Spanish *caballo* would be `r round(items$freq_zipf[items$item=="cat_cavall"]*0.20, 2)`:

$$
\begin{aligned}
\textbf{Exposure}_{\textbf{cavall, Catalan}} &= \textbf{Frequency}_{cavall} \times \textbf{DoE}_{Catalan} = 6.14 \times 0.8 = 4.91 \\
\textbf{Exposure}_{\textbf{caballo, Spanish}} &= \textbf{Frequency}_{caballo} \times \textbf{DoE}_{Spanish} = 6.14 \times 0.2 = 1.23
\end{aligned}
$$

```{r}
#| label: fig-exposure-distribution
df %>% 
    mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
    ggplot(aes(exposure, fill = language)) +
    facet_wrap(~language) +
    geom_histogram(bins = 10, colour = "white", size = 1, show.legend = FALSE) +
    labs(x = "Exposure (Frequency \u00d7 DoE)", y = "Number of observations") +

    df %>% 
    mutate(language = ifelse(grepl("cat_", item), "Catalan", "Spanish")) %>% 
    ggplot(aes(x = freq, y = exposure, colour = language)) +
    facet_wrap(~language) + 
    geom_point(size = 0.1, alpha = 0.5, show.legend = FALSE) +
    labs(x = "Frequency (Zipf score)", y = "Exposure") +
    plot_layout(ncol = 1)
```

## Data analysis

### Model comparison

We initially fitted a base model (`model_fit_0`) that only included the predictors $Age$ (`age_std`), $Phonemes$ (`n_phon_std`), along with random intercepts and slopes of all predictors by $Participant$ (`id`) and $Item$ (`te`), and all correlation parameters. We then expanded this model by adding the main effect of $Exposure$ (`exposure`) (`model_fit_1`), and then $Levenshtein$ (`lv_std`) (`model_fit_2`) and their corresponding random slopes by $Participant$. Then, we added the $Exposure \times Levenshtein$ (`exposure_std:lv_std`) two-way interaction (`model_fit_3`) and its random slope by $Participant$. Finally, we added the $Age \times Exposure \times Levenshtein$ (`age_std:exposure_std:lv_std`) three-way interaction with its random slope by $Participant$, along with a random slope for the $Age \times Exposure$ (`age_std:exposure_std`) two-way interaction.

### Model equation

::: column-body-outset
$$\begin{aligned}
g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
\end{aligned}$$
:::

$s$ is the s-th response category, $a_1 < a_2 < a_3$ are ordered response categories, $\mathrm{Pr}(y_i \le a_3) = 1$ and $\tau_s$ are threshold parameters where $\tau_1 < \tau_2 < \tau_3$.

::: column-body-outset

$$\begin{aligned}
\nu_{ij} = & \beta_0 +\beta_1\mathrm{Age}_{i} + \beta_2\mathrm{Phon}_{j} + \beta_3\mathrm{Exposure}_{ij} + \beta_4 \mathrm{Levenshtein}_{j} + & [\mbox{Main effects}]\\
&\beta_5\mathrm{(Age_{i} \cdot \mathrm{Exposure}_{ij}}) + \beta_6\mathrm{(\mathrm{Age}_{i} \cdot \mathrm{Levenshtein}_{j}}) +  \beta_7\mathrm{(Exposure_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Two-way interactions}]\\
&\beta_8{(\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij} \cdot \mathrm{Levenshtein}_{j}}) + & [\mbox{Three-way interaction}]\\

&\theta_{0j}^{(\mathrm{TE})} + \theta_{1j}^{(\mathrm{TE})}\mathrm{Age}_{i} + \theta_{2j}^{(\mathrm{TE})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{TE})} \mathrm{Exposure}_{ij} +  \theta_{4j}^{(\mathrm{TE})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) + & [\mbox{Item-level effects}]\\

&\theta_{0j}^{(\mathrm{ID})} + \theta_{1j}^{(\mathrm{ID})}\mathrm{Age}_{i} +   \theta_{2j}^{(\mathrm{ID})} \mathrm{Phon}_{j} + \theta_{3j}^{(\mathrm{ID})} \mathrm{Exposure}_{ij} + \theta_{4j}^{(\mathrm{ID})} \mathrm{Levenshtein}_{j} + \theta_{5j}^{(\mathrm{ID})} (\mathrm{Age}_{i} \cdot \mathrm{Exposure}_{ij}) & [\mbox{Participant-level effects}]
\end{aligned}$$
:::

### Model prior

::: column-body-outset
$$\begin{aligned}\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]\end{aligned}$$
:::

In addition, we specified a $LKJ$ prior--$\sim LKJ(2)$ for the correlations between group-level coefficients.

#### Prior predictions

::: {.panel-tabset}

##### Prior-predicted mean

```{r}
#| label: fig-prior-mean
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    exposure_std = 0,
    lv_std = 0
)

m <- add_epred_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA) %>% 
    filter(.category != "No") %>% 
    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(
        c(Understands, `Understands and Says`),
        names_to = ".category",
        values_to = ".epred"
    ) %>% 
    mutate(
        .category = factor(
            .category,
            levels = c("Understands and Says", "Understands"),
            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
        )
    )

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4),
        sec.axis = dup_axis(
            name = "Age (years; months)",
            labels = months_to_years(
                rescale_variable(
                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
                    mean(df$age),
                    sd(df$age)
                )
            )
        )
    ) +
    scale_fill_brewer(palette = "Reds", labels = c("95%", "89%", "67%", "50%")) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

##### Prior-predictions

```{r}
#| label: fig-prior-predictions
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    exposure_std = 0,
    lv_std = 0
)

m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4),
        sec.axis = dup_axis(
            name = "Age (years; months)",
            labels = months_to_years(
                rescale_variable(
                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
                    mean(df$age),
                    sd(df$age)
                )
            )
        )
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

### Model settings

-   OS: `r sessionInfo()$running`, `r sessionInfo()$platform`
-   R version: `r sessionInfo()$R.version$version.string`
-   Algorithm: Hamiltonian Montecarlo (No U-turn Sampler, *NUTS*)
-   Engine: Stan (brms interface, `r sessionInfo()$otherPkgs$brms$Version`)
-   brms backend: CmdStanR (cmdstanr `r sessionInfo()$loadedOnly$cmdstanr$Version`)
-   Chains: `r dim(model_fit_4$fit)[2]`
-   Cores: `r dim(model_fit_4$fit)[2]`
-   Iterations: `r format(dim(model_fit_4$fit)[1], big.mark = ",")` (`r format(dim(model_fit_4$fit)[1]/2, big.mark = ",")`)

More details in Appendix: Session Info.

### R code (`brms`)

::: panel-tabset

#### Model 0

```{r}
print(model_fit_0$formula)
```

#### Model 1

```{r}
print(model_fit_1$formula)
```

#### Model 2

```{r}
print(model_fit_2$formula)
```

#### Model 3

```{r}
print(model_fit_3$formula)
```

#### Model 4

```{r}
print(model_fit_4$formula)
```
:::

# Results

## Model selection

We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) via the `loo` R package. LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set. Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model as:

@tbl-results-loos shows the outputs of the *LOO-CV*. The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference. This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.

```{r}
#| label: tbl-results-loos
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
model_loos %>% 
    loo_compare() %>% 
    as.data.frame() %>%   
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    gt() %>% 
    tab_header(
        title = "",
        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*). The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_3`)")
    ) %>% 
    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
    fmt_number(2:9) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("*SE*"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*")
    ) %>% 
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_title(groups = "subtitle")
    )
```

### Leave-one-out diagnostics (Pareto-*k* diagnostics)

::: panel-tabset

#### Model 0

```{r}
#| label: fig-loo-diagnostics-0
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 0"
get_loo_plot <- function(x){
    x$pointwise[,5] %>% 
        data.frame(x = 1:length(.), y = .) %>% 
        ggplot(aes(x, y)) +
        geom_point(
            size = 0.5, alpha = 0.1, colour = pal_d3()(3)[2], 
            shape = 1
        ) +
        geom_hline(yintercept = 0.5, linetype = "dotted") +
        labs(
            x = "Data point",
            y = "Pareto-k value",
            title = "PSIS diagnostic plot"
        ) +
        scale_x_continuous(breaks = seq(0, length(x$pointwise[,5]), 50000)) +
        scale_y_continuous(
            limits = c(-0.6, 0.6)
        )
}

get_loo_plot(model_loos$model_fit_0)
```

#### Model 1

```{r}
#| label: fig-loo-diagnostics-1
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
get_loo_plot(model_loos$model_fit_1)
```

#### Model 2

```{r}
#| label: fig-loo-diagnostics-2
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 2"
get_loo_plot(model_loos$model_fit_2)
```

#### Model 3

```{r}
#| label: fig-loo-diagnostics-3
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 3"
get_loo_plot(model_loos$model_fit_3)
```

#### Model 4

```{r}
#| label: fig-loo-diagnostics-4
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 4"
get_loo_plot(model_loos$model_fit_4)
```

:::

## Fixed effects

```{r}
#| label: results-posterior-fixed
#| echo: true
#| message: false
#| warning: false
post_draws <- gather_draws(model_fit_4, `b_.*`, regex = TRUE)

# tidy predictor names
str_repl <- c(
    "b_Intercept[1]" = paste0("Intercept (Comprehension and Production at ",  round(mean(df$age, 2)), " months)"),
    "b_Intercept[2]" = paste0("Intercept (Comprehension at ", round(mean(df$age, 2)), " months)"),
    "b_age_std" = paste0("Age (+1 SD, ",  round(sd(df$age), 2), " months)"),
    "b_n_phon_std" = paste0("Phonemes (+1 SD, ",  round(sd(df$n_phon), 2), " phonemes)"),
    "b_exposure_std" = paste0("Exposure (+1 SD, ",  percent(sd(df$exposure)), ")"),
    "b_lv_std" = paste0("Levenshtein (+1 SD, ",  percent(sd(df$lv)), ")"),
    "b_exposure_std:lv_std" = "Exposure \u00d7 Levenshtein",
    "b_age_std:exposure_std" = "Age \u00d7 Exposure",
    "b_age_std:lv_std" = "Age \u00d7 Levenshtein",
    "b_age_std:exposure_std:lv_std" = "Age \u00d7 Exposure \u00d7 Levenshtein"
)

```

```{r}
#| label: tbl-results-fixed
#| tbl-cap: "Summary of the estimated posterior distribution of fixed regression coefficients"
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
intercept_footnote <- paste0(
    "Probability of comprehension or production at the grand age of the sample (", 
    round(mean(df$age), 2), " months)"
)

# summarise posterior draws
posterior_description %>% 
    mutate(
        parameter = factor(
            parameter, 
            levels = names(str_repl), 
            labels = str_repl
        )
    ) %>%  
    select(parameter, median, ci_high, ci_low, rope_percentage, rhat, ess) %>% 
    gt(rowname_col = "parameter") %>% 
    fmt_number(c(1:4, 6:7)) %>% 
    fmt_percent(5) %>% 
    cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>% 
    cols_label(
        median = "Median",
        rhat = "R-hat",
        ess = "ESS",
        ci_low = md("95% *HDI*"),
        rope_percentage = md("*ROPE* prob.")
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = 1:7)
    )
post_draws_list <- split(posterior_description, posterior_description$parameter)

```

We specified a region of practical equivalence (*ROPE*) around the values of each regression coefficient that we considered equivalent to zero [@kruschke2018bayesian]. We decided to draw this interval around zero so that it covered values smaller than those considered as negligible by similar previous studies: after consulting the regression tables of [ADD STUDIES], we selected the largest regression coefficient that was interpreted as conclusive evidence of the presence of an effect, as indicated by the authors, and transformed it to the probability scale: 3%. consequently, we defined our *ROPE* from -3% to +3%. We report the mean of each coefficient's marginal posterior distribution along with their 95% highest density intervals (*HDI*). This interval indicates the narrowest range of values that contain 95% of the samples of the posterior distribution, and therefore the true value of the coefficient with 95% confidence, given our data. Following [@kruschke2018bayesian], we considered as relevant those regression coefficients whose 95% *HDI* did not overlap with the -3%--3% *ROPE* interval we previously defined. For those coefficients that did not exclude completely the *ROPE* we also report a Bayes Factor indicating how many times is such coefficient more likely to be bigger than the upper limit of the *ROPE*, or smaller than the lower limit of the *ROPE*, given our data.

For interpretability, we report the estimated regression coefficients transformed from the probability scale. The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales related non-linearly. This means that one logit difference is not necessarily translated to a unique value in the probability scale. For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 34 to 36 months. The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients  in the probability scale is not trivial. We followed the suggestion by @gelman2020regression and report the maximum value of such coefficient, which corresponds to the linear growth (i.e. derivative) of the logistic curve at the age at which most participants were acquiring a given word. This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].

@fig-results-fixed shows the outputs of the model. **Age** was the predictor with the largest effect on the probability of acquisition ($\beta$ = `r printnum(post_draws_list[["b_age_std"]]$median, digits = 2)`, 95% *HDI* = [`r printnum(post_draws_list[["b_age_std"]]$ci_low, digits = 2)`, `r printnum(post_draws_list[["b_age_std"]]$ci_high, digits = 2)`]): at the steepest point of the acquisition trajectory of an average translation equivalent, the probability of acquisition increased `r percent(post_draws_list[["b_age_std"]]$median/(4*sd(df$age)), accuracy = 0.01)` every month.

The 95% *HDI* of the regression coefficient of the **number of phonemes** did exclude the *ROPE*, although not by much: `r percent(post_draws_list[["b_n_phon_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*. This suggests that the number of phonemes in the word form had a small effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_n_phon_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_n_phon_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_n_phon_std"]]$ci_high, accuracy = 0.01)`]). For every increase in phoneme length, the probability of acquisition decreased in `r percent(post_draws_list[["b_n_phon_std"]]$median/(4*sd(df$n_phon)), accuracy = 0.01)`.

**Exposure** ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$ci_high, accuracy = 0.01)`]). All of the posterior samples of his regression excluded the *ROPE*. The impact of this predictor on the probability of acquisition was positive: for every SD increase in exposure rate, the participant was `r percent(post_draws_list[["b_exposure_std"]]$median/(sd(df$b_exposure_std)), accuracy = 0.01)` more likely to acquire it.

The effect of **phonological similarity** by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$ci_high, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$rope_percentage, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.

However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to the word. Follow up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to the language was low (e.g., 10%), phonological similarity increased the probability of acquisition substantially. This effect was negligible when exposure to both languages was balanced (e.g. 50-50%). Finally, when exposure to the language was high (e.g., 90%), this effect was very close to zero, although with different direction, with phonological less similar translation equivalents being more likely to be acquired than phonologically more similar ones.

[TO DO: emmeans tests]

The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:doe_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$ci_high, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$rope_percentage, accuracy = 0.01)` of its posterior samples fell within the *ROPE*. This shows that the effect of exposure on the speed at which participants acquired word across ages was inconclusive. Give our data, it is difficult to tell whether this *Exposure* increased or decreased the slope of the words' acquisition trajectories.

The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$ci_high, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages.

Finally, the 95% *HDI* of the regression coefficient of the $Age \times Exposure \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_low, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$ci_high, accuracy = 0.01)`]), suggesting that the impact of exposure to the language on the effect of phonological similarity did not differ substantially across ages.

```{r}
#| label: fig-results-fixed
#| echo: true
#| message: false
#| warning: false
#| fig-width: 9
#| fig-heigh: 7
#| fig-cap: "Marginal posterior distribution of the regression coefficients of the fixed effects in the extended model. Distributions show the estimated likelihood density of each value in the parameter space (X-axis) of each coefficient (Y-axis). Intervals show the 95% highest density interval (*HDI*) of each distribution. This interval is the narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data. the mean and HDI limits are indicated below each distribution. Coefficients were transformed from the logit scale to the probability scale for interpretability. We used the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in this variable, i.e. the derivative of the logistic function in its mid-point."
post <- post_draws %>% 
    mutate(
        .variable_name = factor(
            .variable, 
            levels = names(str_repl),
            labels = str_repl,
            ordered = TRUE
        )
    )

post %>%
    filter(!grepl("Intercept|sd", .variable)) %>%
    ggplot(aes(.value, fct_rev(.variable_name))) +
    annotate(
        geom = "rect",
        ymin = -Inf,
        ymax = Inf,
        xmin = rope_interval["lower"],
        xmax = rope_interval["upper"],
        colour = NA,
        alpha = 0.2,
        fill = pal_d3()(3)[2],
    ) +
    geom_vline(
        xintercept = 0,
        size = 1,
        colour = pal_d3()(3)[2]
    ) +
    stat_slab(
        aes(
            fill = stat(abs(x) < rope_interval["upper"]),
            colour = stat(abs(x) < rope_interval["upper"]),
        ),
        size = 1,
        # fill = pal_d3()(2)[2],
        position = position_nudge(y = 0.15)
    ) +
    geom_errorbar(
        data = post %>% 
            filter(!grepl("Intercept|sd", .variable)) %>%
            median_hdi(
                .exclude = c(
                    ".chain", 
                    ".iteration",
                    ".draw",
                    ".row", 
                    ".variable_name"
                )
            ) %>% 
            mutate(
                .variable_name = factor(
                    .variable, 
                    levels = names(str_repl),
                    labels = str_repl,
                    ordered = TRUE
                )
            ),
        aes(xmin = .lower, xmax = .upper, x = .value),
        width = 0.15
    ) +
    geom_text(
        data = post %>% 
            filter(!grepl("Intercept|sd", .variable)) %>%
            mean_hdi(
                .exclude = c(".chain", ".iteration", ".draw", ".row", ".variable_name")
            ) %>% 
            mutate(
                .variable_name = factor(
                    .variable, 
                    levels = names(str_repl),
                    labels = str_repl,
                    ordered = TRUE
                )
            ),
        aes(
            label = paste0(
                printnum(.value, digits = 2),
                " [", printnum(.lower, digits = 2), ", ",
                printnum(.upper, digits = 2), "]"
            )
        ),
        position = position_nudge(y = -0.2),
        size = 3
    ) +
    # annotate(
    #     geom = "text",
    #     label = "ROPE",
    #     x = 0.3,
    #     y = 10,
    #     vjust = 1,
    #     size = 5,
    #     colour = pal_d3()(2)[2]
    # ) +
    # annotate(
    #     geom = "curve",
#     arrow = arrow(length = unit(0.2, "cm")),
#     x = 0.3,
#     y = 9.5,
#     xend = 0.1,
#     yend = 9,
#     size = 1,
#     curvature = -0.25,
#     colour = pal_d3()(2)[2]
# ) +
labs(
    x = "Logistic regression coefficient estimate",
    y = "Variable", 
    fill = "Overlaps with ROPE",
    colour = "Overlaps with ROPE"
) +
    scale_x_continuous(breaks = seq(-0.5, 2, 0.2)) +
    scale_fill_manual(
        values = c("dodgerblue", "grey"),
        labels = c("No", "Yes")
    ) +
    scale_colour_manual(
        values = c("dodgerblue", "grey"),
        labels = c("No", "Yes")
    ) +
    theme(
        legend.position = "top",
        legend.justification = c(1, 1),
        axis.title.y = element_blank(),
        panel.grid.major.x = element_line(
            colour = "grey85",
            linetype = "dotted"
        ),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

## Marginal means

```{r}
#| label: fig-results-marginal
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected posterior predictions. Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient. These values correspond to the 4,000 samples of the posterior distribution that were drawn during the MCMC estimation of the model. Therefore, we generated 4,000 predictions for a series of combination of levels of interest. Each line and interval and in the figure summarises the mean and 95% *HDI* of each combination of levels. The X-axis indicates the age (in months) for which the prediction is generated. The Y-axis indicates the predicted probability of acquisition (*Comprehension* or *Comprehension and Production*). Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents). Finally, predictions are presented separately for different degrees of exposure (10%): little exposure to the language (10% DoE, traditionally classified as monolingual), balanced exposure to both languages (50%, traditionally classified as bilingual), and high exposure to the language (90%, traditionally classified as monolingual). Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom."
nd <- expand.grid(
    n_phon_std = 0,
    freq_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    exposure_std = c(min(df$exposure_std, na.rm = TRUE), 0, max(df$exposure_std, na.rm = TRUE)),
    lv_std = (c(0, 1)-mean(df$lv))/(sd(df$lv))
)

m <- add_epred_draws(nd, model_fit_4, ndraws = 50, re_formula = NA) %>% 
    mutate(
        freq_std = as.factor(paste0("Frequency = ", freq_std, " SD")),
        exposure_std = factor(
            exposure_std, 
            levels = unique(nd$exposure_std), 
            labels = c("Minimum exposure", "Mean exposure", "Maximum exposure")
        ),
        lv_std = factor(lv_std, levels = unique(nd$lv_std), labels = paste0(c(0, 100), "% similarity")
        )
    ) %>% 
    filter(.category != "No") %>% 
    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(Understands, `Understands and Says`), names_to = ".category", values_to = ".epred") %>% 
    mutate(
        .category = factor(
            .category,
            levels = c("Understands", "Understands and Says"),
            labels = c("Comprehension\n(Understands)", "Production\n(Understands and Says")
        )
    )

d <- df %>%
    mutate(
        lv_std = case_when(
            between(lv, 0.00, 0.10) ~ "0% similarity",
            between(lv, 0.90, 1.00) ~ "100% similarity"
        ),
        exposure_std = case_when(
            between(exposure_std, min(df$exposure_std, na.rm = TRUE), min(df$exposure_std, na.rm = TRUE)+0.5) ~ "Minimum exposure",
            between(exposure_std, mean(df$exposure_std, na.rm = TRUE)-0.25, mean(df$exposure_std, na.rm = TRUE)+0.25) ~ "Mean exposure",
            between(exposure_std, max(df$exposure_std, na.rm = TRUE)-0.5, max(df$exposure_std, na.rm = TRUE)) ~ "Maximum exposure"
        ),
        age_std = ((floor(age)-mean(df$age))/(sd(df$age))) %>% 
            cut(
                .,
                breaks = ((seq(10, 38, 2)-mean(df$age))/(sd(df$age))),
                labels = ((seq(10, 36, 2)-mean(df$age))/(sd(df$age))),
                include.lowest = TRUE
            ) %>% 
            as.character() %>% 
            as.numeric(),
        Understands = response %in% c("Understands", "Understands and Says"),
        `Understands and Says` = response %in% "Understands and Says"
    ) %>% 
    drop_na(exposure_std, lv_std) %>% 
    group_by(te, age_std, exposure_std, lv_std) %>% 
    summarise(
        Understands = mean(Understands),
        `Understands and Says` = mean(`Understands and Says`),
        .groups = "drop"
    ) %>% 
    pivot_longer(
        c(Understands, `Understands and Says`),
        names_to = ".category",
        values_to = ".epred"
    ) %>% 
    mutate(
        .category = factor(
            .category,
            levels = c("Understands", "Understands and Says"),
            labels = c("Comprehension\n(Understands)", "Production\n(Understands and Says")
        )
    )

m %>% 
    ggplot() +
    aes(
        x = age_std, 
        y = .epred, 
        colour = lv_std, 
        fill = lv_std,
    ) +
    facet_grid(.category~exposure_std) +
    # stat_lineribbon(
    #     size = 1,
    #     alpha = 0.5,
    #     .width = 0.95, 
    #     colour = NA,
    #     point_interval = mean_hdi
    # ) +
    geom_line(aes(group = interaction(.draw, lv_std)), alpha = 0.25) +
    stat_summary(fun = "mean", geom = "line", size = 1) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_color_d3() +
    labs(
        x = "Age (months)", 
        y = "Posterior probability of acquisition\n(Linear prediction)",
        colour = "Phonological similarity (Levenshtein distance)", 
        fill = "Phonological similarity (Levenshtein distance)"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

## Random effects

### Participant-level effects

::: panel-tabset

#### Intercept

```{r}
#| label: fig-results-random-id-intercept
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior probability of acquisition by participant"
re_id <- ranef(model_fit_4)$id[,,1] %>% 
    as_tibble() %>% 
    rownames_to_column("id") %>% 
    left_join(distinct(model_fit_4$data, id)) %>% 
    clean_names() %>% 
    mutate(
        across(
            .cols = where(is.numeric),
            .fns = list(
                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
            ), .names = "{.col}__{.fn}"
        )
    ) %>% 
    select(-c(estimate:q97_5)) %>% 
    pivot_longer(
        where(is.numeric), 
        names_to = c(".value", "type"),
        names_sep = "__"
    )

re_id %>% 
    ggplot() +
    aes(
        x = estimate, 
        y = reorder(id, estimate), 
        xmin = q2_5, 
        xmax = q97_5,
        fill = type,
        colour = type
    ) +
    facet_wrap(~type) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior probability of acquisition",
        y = "Participant", 
        colour = "Type"
    ) +
    scale_x_continuous(limits = c(0, 1), labels = percent) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Age slope

```{r}
#| label: fig-results-random-id-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_age <- ranef(model_fit_4)$id[,,"age_std"]

re_age %>% 
    as_tibble() %>% 
    clean_names() %>% 
    mutate(id = unique(model_fit_4$data$id)) %>% 
    relocate(id) %>% 
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant", 
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### # Phonemes slope

```{r}
#| label: fig-results-random-id-nphon
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the number of phonemes by participant"
re_n_phon <- ranef(model_fit_4)$id[,,"n_phon_std"]

re_n_phon %>% 
    as_tibble() %>% 
    clean_names() %>% 
    mutate(id = unique(model_fit_4$data$id)) %>% 
    relocate(id) %>% 
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant", 
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Exposure slope

```{r}
#| label: fig-results-random-id-exposure
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of lexical frequency by participant"
re_freq <- ranef(model_fit_4)$id[,,"exposure_std"]

re_freq %>% 
    as_tibble() %>% 
    clean_names() %>% 
    mutate(id = unique(model_fit_4$data$id)) %>% 
    relocate(id) %>% 
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant", 
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Levenshtein slope

```{r}
#| label: fig-results-random-id-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of phonological similarity by participant"
re_lv <- ranef(model_fit_4)$id[,,"lv_std"]

re_lv %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(id = unique(model_fit_4$data$id)) %>%
    relocate(id) %>%
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant",
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Doe-by-Levenshtein slope

```{r}
# | label: fig-results-random-id-exposure-lv
# | echo: true
# | message: false
# | warning: false
# | fig-width: 7
# | fig-cap: "Posterior random slope of age by participant"
re_exposure_lv <- ranef(model_fit_4)$id[,,"exposure_std:lv_std"]

re_lv %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(id = unique(model_fit_4$data$id)) %>%
    relocate(id) %>%
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant",
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Age-by-DoE slope

```{r}
#| label: fig-results-random-id-age-doe
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-language degree of exposure interaction by participant"
re_age_exposure <- ranef(model_fit_4)$id[,,"age_std:exposure_std"]

re_age_exposure %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(id = unique(model_fit_4$data$id)) %>%
    relocate(id) %>%
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant",
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Age-by-Levenshtein slope

```{r}
#| label: fig-results-random-id-age-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-phonological similarity interaction by participant"
re_age_lv <- ranef(model_fit_4)$id[,,"age_std:lv_std"]

re_age_lv %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(id = unique(model_fit_4$data$id)) %>%
    relocate(id) %>%
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant",
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Age-by-DoE-by-Levenshtein slope

```{r}
#| label: fig-results-random-id-age-exposure-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the age-by-exposure-by-phonological similarity interaction by participant"
re_age_exposure_lv <- ranef(model_fit_4)$id[,,"age_std:exposure_std:lv_std"]

re_age_exposure_lv %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(id = unique(model_fit_3$data$id)) %>%
    relocate(id) %>%
    ggplot() +
    aes(
        x = estimate,
        y = reorder(id, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant",
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Correlations

```{r}
#| label: fig-results-random-id-correlations
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior correlations between random effects by participant"
posterior_description_re %>% 
    select(parameter, median) %>% 
    filter(str_detect(parameter, "cor_id__")) %>% 
    mutate(parameter = str_remove_all(parameter, "cor_id__")) %>%  
    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
    mutate(
        across(
            c(var_1, var_2), 
            ~factor(
                .,
                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std", "exposure_std:lv_std", "age_std:exposure_std:lv_std"),
                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein", "Exposure \u00d7 Levenshtein", "Age \u00d7 Exposure \u00d7 Levenshtein")
            )
        )
    ) %>%
    filter(var_1 != var_2) %>% 
    ggplot(aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median))) +
    geom_point(shape = 15, size = 20) +
    geom_text(colour = "black") +
    scale_color_distiller(palette = "RdYlBu") +
    labs(colour = "Correlation", size = "Correlation") +
    theme(
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 35, hjust = 1),
        panel.grid = element_blank()
    )
```
:::

### Item-level effects

::: panel-tabset

#### Intercept

```{r}
#| label: fig-results-random-te
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior probability of acquisition by translation equivalent"
re_te <- ranef(model_fit_4)$te[,,"Intercept"] %>% 
    as_tibble() %>% 
    rownames_to_column("te") %>% 
    clean_names() %>% 
    mutate(
        across(
            .cols = where(is.numeric),
            .fns = list(
                Comprehension = ~inv_logit_scaled(. + fixef(model_fit_4)[1]),
                Production = ~inv_logit_scaled(. + fixef(model_fit_4)[2])
            ), .names = "{.col}__{.fn}"
        )
    ) %>% 
    select(-c(estimate:q97_5)) %>% 
    pivot_longer(
        where(is.numeric), 
        names_to = c(".value", "type"),
        names_sep = "__"
    ) %>% 
    mutate(te = as.numeric(te))

re_te %>% 
    ggplot() +
    aes(
        x = estimate, 
        y = reorder(te, estimate), 
        xmin = q2_5,
        xmax = q97_5,
        colour = type,
        fill = type
    ) +
    facet_wrap(~type) +
    geom_ribbon(aes(group = 1), alpha = 0.5, colour = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0.5, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior probability of acquisition",
        y = "TE", 
        colour = "Type"
    ) +
    scale_x_continuous(limits = c(0, 1), labels = percent) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Age slope

```{r}
#| label: fig-results-random-te-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by translation equivalent"
re_age <- ranef(model_fit_4)$te[,,"age_std"]

re_age %>% 
    as_tibble() %>% 
    clean_names() %>% 
    mutate(te = unique(model_fit_4$data$te)) %>% 
    relocate(te) %>% 
    ggplot() +
    aes(
        x = estimate,
        y = reorder(te, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant", 
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### # Phonemes slope

```{r}
#| label: fig-results-random-te-nphon
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of the number of phonemes by translation equivalent"
re_n_phon <- ranef(model_fit_3)$te[,,"n_phon_std"]

re_n_phon %>% 
    as_tibble() %>% 
    clean_names() %>% 
    mutate(te = unique(model_fit_3$data$te)) %>% 
    relocate(te) %>% 
    ggplot() +
    aes(
        x = estimate,
        y = reorder(te, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant", 
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Exosure slope

```{r}
#| label: fig-results-random-te-exposure
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_doe <- ranef(model_fit_4)$te[,,"exposure_std"]

re_doe %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(te = unique(model_fit_4$data$te)) %>%
    relocate(te) %>%
    ggplot() +
    aes(
        x = estimate,
        y = reorder(te, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant",
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Age-by-exposure slope

```{r}
#| label: fig-results-random-te-age-exposure
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior random slope of age by participant"
re_age_doe <- ranef(model_fit_4)$te[,,"age_std:exposure_std"]

re_age_doe %>% 
    as_tibble() %>% 
    clean_names() %>% 
    mutate(te = unique(model_fit_4$data$te)) %>% 
    relocate(te) %>% 
    ggplot() +
    aes(
        x = estimate,
        y = reorder(te, estimate),
        xmin = q2_5,
        xmax = q97_5
    ) +
    geom_ribbon(aes(group = 1), alpha = 0.5, color = NA) +
    geom_line(aes(group = 1), size = 1) +
    geom_vline(xintercept = 0, colour = "black", linetype = "dotted") +
    labs(
        x = "Posterior random slope estimate",
        y = "Participant", 
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_line(linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

#### Correlations

```{r}
#| label: fig-results-random-te-correlations
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Posterior correlations between random effects by translation equivalent"
posterior_description_re %>% 
    select(parameter, median) %>% 
    filter(str_detect(parameter, "cor_te__")) %>% 
    mutate(parameter = str_remove_all(parameter, "cor_te__")) %>%  
    separate(parameter, c("var_1", "var_2"), sep = "__") %>%
    mutate(
        across(
            c(var_1, var_2), 
            ~factor(
                .,
                levels = c("Intercept", "age_std", "n_phon_std", "exposure_std", "lv_std", "age_std:exposure_std", "age_std:lv_std"),
                labels = c("Intercept", "Age", "Phonemes", "Exposure", "Levenshtein", "Age \u00d7 Exposure", "Age \u00d7 Levenshtein")
            )
        )
    ) %>%
    filter(var_1 != var_2) %>% 
    ggplot() +
    aes(x = var_1, y = fct_rev(var_2), colour = median, label = printnum(median)) +
    geom_point(shape = 15, size = 40) +
    geom_text(colour = "black") +
    scale_color_distiller(palette = "RdYlBu") +
    labs(
        colour = "Correlation",
        size = "Correlation"
    ) +
    theme(
        axis.title = element_blank(),
        panel.grid = element_blank()
    )
```
:::

## Model diagnostics

### Traceplots

```{r}
# | label: fig-diagnostics-traceplots
# | echo: true
# | message: false
# | warning: false
# | fig-width: 12
gather_draws(model_fit_4, `b_.*`, `sd_te__.*`, regex = TRUE) %>%
    mutate(.chain = paste("Chain ", .chain)) %>%
    ggplot(aes(.iteration, .value, colour = .chain)) +
    facet_wrap(~.variable, scales = "free_y") +
    annotate(
        geom = "rect",
        colour = NA,
        xmin = 0,
        xmax = dim(model_fit_4$fit)[1]/2,
        ymin = -Inf,
        ymax = Inf,
        alpha = 0.25
    ) +
    geom_line() +
    labs(
        x = "Iteration",
        y = "Sample value",
        colour = "Chain"
    ) +
    scale_color_d3() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = 7)
    )
```

### Gelman-Rubin diagnostic (R-hat)

```{r}
#| label: fig-diagnostics-rhat
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| eval: false
mcmc_rhat_data(model_rhats$model_fit_4) %>% 
    ggplot() +
    aes(x = value) +
    geom_histogram(
        bins = 100, 
        na.rm = TRUE, 
        fill = pal_d3()(1)[1], 
        colour = "white"
    ) +
    geom_rug(alpha = 0.01, colour = pal_d3()(1)[1]) +
    labs(
        x = "R-hat",
        y = "Number of samples",
        colour = "Description",
        fill = "Description"
    ) +
    theme(
        legend.position = "top"
    )
```

### Effective sample size

```{r}
#| label: fig-diagnostics-neff
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| eval: false
mcmc_neff_data(model_neffs$model_fit_4) %>% 
    ggplot() +
    aes(x = value, fill = description, colour = description) +
    geom_histogram(
        bins = 100, 
        na.rm = TRUE, 
        colour = "white"
    ) +
    geom_rug(alpha = 0.01) +
    geom_vline(xintercept = 1, colour = "black", size = 0.75) +
    labs(
        x = "Effective sample size ratio",
        y = "Number of samples",
        colour = "Rating",
        fill = "Rating"
    ) +
    theme(
        legend.position = "top"
    )
```

# Discussion

## Summary of the study

## Summary of the results

## Results in contrast with the literature

## Mechanistic explanations

## Limitations

## Further steps

# Appendix

## Session info

```{r}
sessionInfo()
```