### 2022-12-09 dag.tex
Update model formula and add DAG

\begin{tikzpicture}

  % nodes
  \node (age) at (0,0) {$Age_{i, j}$};
  \node (phonemes) at (3,4) {$Phonemes_{i, j}$};
  \node (exposure) at (3,2) {$Exposure_{i, j}$};
  \node (cognateness) at (3,-2) {$Cognateness_{i}$};
  \node [ellipse, draw](p_acquisition) at (7, 0) {$P(Acquisition)$};
  \node (response) at (11, 0) {$Response$};
  % arrows
  \draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (exposure);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (age) -- (cognateness);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (exposure) -- (cognateness);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (phonemes) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (cognateness) -- (p_acquisition);
  \draw [->, shorten >= 4pt, shorten <= 4pt] (p_acquisition) -- (response);

\end{tikzpicture}