### 2023-06-28 manuscript/appendix.qmd
Update PDF format

---
title: "Appendix"
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

fig-align: center
fig-dpi: 800
echo: false
warning: false
message: false

format:
  pdf:
    documentclass: article
    papersize: letter
    fontsize: 12pt
    linestretch: 1.5
    header-includes:
    - \usepackage{tipa}
    geometry:
    - top=1in
    - bottom=1in
    - left=1in
    - right=1in
    colorlinks: true
    linkcolor: DodgerBlue
    urlcolor: DodgerBlue
    citecolor: DodgerBlue
    link-citations: true
    biblio-config: false
    block-headings: false
    indent: true
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

library(patchwork)
library(gt)
library(ggplot2)

targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

targets::tar_load_globals()

targets::tar_load(c(bvq_data, childes, items, 
                    participants, responses))

targets::tar_load(c(model_fit, model_summary, model_epreds,
                    model_convergence, model_fit_prior, 
                    model_ppcs, posterior_doe_summary,
                    posterior_doe_draws))

targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))

# process objects
items <- bvq_data$pool  |>  
    mutate(item = str_remove(item, "cat_|spa_")) |> 
    select(item, language, semantic_category, class) |> 
    right_join(items, by = c("item", "language"))

# set ggplot theme and colour palette ------------------------------------------

my_theme <- function() {
    theme_ggdist() +
        theme(panel.grid = element_line(colour = "grey",
                                        linetype = "dotted"),
              axis.line = element_line(colour = "black"),
              text = element_text(size = 12, colour = "black"),
              axis.text = element_text(colour = "black"))
}

theme_set(my_theme())

bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))

set.seed(888)

options(knitr.graphics.error = FALSE)
```

## Appendix A: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_model-doe.qmd >}}

{{< pagebreak >}}

## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_syllable-frequency.qmd >}}

{{< pagebreak >}}