### 2023-07-06 docs/index.qmd
Fix appendix

---
title: "Cognate beginnings to bilingual lexical acquisition"
subtitle: "Lab notes"
short-title: "Cognate beginnings"
author:
  - name: Gonzalo Garcia-Castro
    corresponding: true
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    affiliations:
      - id: cbc
        name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias Fargas 25-27
        city: Barcelona
        region: Spain
        postal-code: "08005"
  - name: Daniela S. Avila-Varela
    orcid: 0000-0002-3518-8117
    affiliations: 
      - ref: cbc
      - id: ul
        name: Universidade de Lisboa
        department: Facultade de Letras
        address: Alameda da Universidade
        city: Lisboa
        region: Portugal
        postal-code: "1600-214"
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations: 
      - id: uam 
        name: Universidad Autónoma de Madrid
        department: Departamento de Psicología
        address: Iván Pavlov, 6
        city: Madrid
        region: Spain
        postal-code: "28049"
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations: 
      - ref: cbc
date: "May 10th, 2023"
code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
keywords:
  - lexical acquisition
  - vocabulary
  - bilingualism
  - item response theory
  - bayesian
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

code-fold: true
warning: false

format:
  html: default

fig-dpi: 1000
---

```{r setup}
#| label: setup
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
thanks <- yaml_metadata$thanks

```

::: callout-tip
Data and materials are available at [OSF](https://osf.io/hy984/), and
code is available at
[GitHub](https://github.com/gongcastro/trajectories).
:::

`r abstract`

`r thanks`

# Links

![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height=100px}

![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height=100px}

![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height=100px}

{{< include ../README.md >}}

# Session info

```{r}
sessionInfo()
```