### 2023-07-23 docs/index.qmd
Simplify dependencies

---
---

```{r setup}
#| label: setup
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

yaml_metadata <- rmarkdown::yaml_front_matter(
    here::here("manuscript", "manuscript.qmd")
)

title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
thanks <- yaml_metadata$thanks
```

::: callout-tip
Data and materials are available at [OSF](https://osf.io/hy984/), and
code is available at
[GitHub](https://github.com/gongcastro/trajectories).
:::

`r abstract`

`r thanks`

# Links

![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height=100px}

![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height=100px}

![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height=100px}

{{< include ../README.md >}}

# Session info

```{r}
sessionInfo()
```