### 2023-07-23 manuscript/_appendix/_model-doe.qmd

As a robustness check, we fit a model similar to the one described in the main manuscript, but including lexical frequency and language degree of exposure as separate predictors, instead of the composite measure *Exposure*. Language degree of exposure (*DoE*) was included in interaction with *Age* and *Cognateness*, while lexical frequency (*Frequency*) was included as a main effect. @tbl-coefs-doe shows a comparison between the posterior distribution of the regression coefficients of both models. Overall, results are equivalent.

```{r tbl-coefs-doe}
#| label: tbl-coefs-doe
#| tbl-cap: "Posterior distribution of regression coefficients of the model including the *Exposure* composite predictor, and of the model including lexical frequency (*Frequency*) and degree of exposure (*DoE*) separately. Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution. *p*(ROPE): overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero. In the right column, we show a graphical representation of the median and 95% HDI of each coefficient, with a reference dotted line indicating the location of zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
tbl_data <- list("Exposure model" = model_summary,
                 "DoE-Frequency model" = posterior_doe_summary) |> 
    purrr::map_df(\(x) mutate(x, across(.variable_name, as.character)),
                  .id = ".model") |> 
    mutate(
        across(c(.value, .lower, .upper),
               function(x) {
                   ifelse(
                       .type=="Intercepts (at 22 months)",
                       plogis(x),
                       x/4
                   )
               }
        )) |>
    filter(.type=="Slopes") |> 
    select(.model, .variable_name, .value, .lower, .upper, .rope) |> 
    mutate(
        across(.value:.upper, plogis),
        .variable_name = gsub("×", '$\\\\times$', .variable_name),
        .hdi = paste0("[", round(.lower, 3),
                      ", ", round(.upper, 3), "]"),
        .rope = sub("^(-?)0.", "\\1.",
                    sprintf("%.3f", .rope))) |> 
    select(.variable_name, .value, .lower, .upper, .hdi, .rope)

tbl_data |> 
    mutate(viz = "") |> 
    select(-c(.lower, .upper)) |> 
    knitr::kable(
        format = "latex", 
        escape = FALSE,
        col.names = c("",
                      "$\\beta$",
                      "95\\% HDI",
                      "$p(\\text{ROPE})$",
                      ""),
        digits = 3,
        booktabs = TRUE,
        align = "lclr",
        linesep = ""
    ) |> 
    kableExtra::pack_rows("Model: Exposure",
                          start_row = 1,
                          end_row = 8) |> 
    kableExtra::pack_rows("Model: Frequency & DoE",
                          start_row = 9,
                          end_row = 17)
```