### 2023-07-23 manuscript/appendix.qmd

---
csl: "_assets/apa7.csl"
bibliography: "_assets/references.bib"
link-citations: true
editor: source

fig-align: center
fig-dpi: 800
echo: false
warning: false
message: false

format:
  arxiv-pdf:
    linenumbers: true
    doublespacing: false
    runninghead: "Cognate beginnings to bilingual lexical acquistion (Appendix)"
    include-in-header: 
      text: |
        \usepackage{tipa}
        \usepackage{booktabs}
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
library(knitr)
library(kableExtra)
library(ggplot2)

targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

targets::tar_load_globals()

targets::tar_load(c(bvq_data, childes, items, model_summary,
                    participants, responses))

targets::tar_load(c(posterior_doe_summary, posterior_doe_draws))

targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables,
                    posterior_syllables_summary))

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(colour = "grey",
                                    linetype = "dotted"),
          axis.line = element_line(colour = "black"),
          text = element_text(size = 12, colour = "black"),
          axis.text = element_text(colour = "black"))

theme_set(my_theme)
```

## Appendix A: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_model-doe.qmd >}}

{{< pagebreak >}}

## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_syllable-frequency.qmd >}}

{{< pagebreak >}}

## References