### 2023-07-25 docs/index.qmd
Update Dockerfile

```{r setup}
#| label: setup
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

yaml_metadata <- rmarkdown::yaml_front_matter(
    here::here("manuscript", "manuscript.qmd")
)

title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
thanks <- yaml_metadata$thanks
```

::: callout-tip
Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
:::

`r abstract`

`r thanks`

# Links

![arXiv](https://info.arxiv.org/brand/images/brand-logo-primary.jpg){height="100px"}

![OSF](https://www.datacc.org/wp-content/uploads/2020/03/osf.png){height="100px"}

![GitHub](https://upload.wikimedia.org/wikipedia/commons/thumb/9/91/Octicons-mark-github.svg/2048px-Octicons-mark-github.svg.png){height="100px"}

{{< include ../README.md >}}

# Data dictionaries

```{r}
#' Create data dictionary for a dataframe and save as CSV in the docs/ folder
#' 
#' @param x Data frame for which to create a data dictionary
data_dict <- function(x) {
    if (!(is.data.frame(x) | tibble::is_tibble(x))) {
        cli::cli_abort("x must be a dataframe or tibble")
    }

    x <- flatten_columns(x)

    nmissing <- function(x) sum(is.na(x))
    nlevels <- function(x) length(unique(x))

    # describe character variables
    x_character <- x |> 
        summarise(across(where(\(x) is.character(x) | is.factor(x)),
                         tibble::lst(nmissing, nlevels),
                         .names = "{.col}.{.fn}"))

    # describe numeric variables
    numeric.fns <- tibble::lst(nmissing, mean, median, sd, min, max)
    x_numeric <- x |> 
        summarise(across(where(is.numeric), 
                         numeric.fns,
                         .names = "{.col}.{.fn}"))

    # describe date variables
    date.fns <- tibble::lst(nmissing, min, max)
    x_date <- x |> 
        summarise(across(where(is.Date), 
                         date.fns,
                         .names = "{.col}.{.fn}"))

    # join
    out <- bind_cols(x_character, x_numeric, x_date) |> 
        pivot_longer(everything(),
                     names_to = c("variable", "measure"),
                     values_to = "value",
                     names_sep = "\\.", ) |> 
        pivot_wider(id_cols = variable,
                    names_from = "measure",
                    values_from = "value",
                    values_fn = \(x) ifelse(is.character(x),
                                            paste0(x, collapse = ","),
                                            unlist(x))) |> 
        mutate(varclass = purrr::map_chr(x, class)) |> 
        relocate(varclass, .after = variable)

    return(out)
}
```

::: {.callout-info}

## Items

```{r items-dict}
tar_load(items)

items_dict <- data_dict(items) |> 
    mutate(comments = c(
        "Concept associated to the word-form, uniquely identifies every translation equivalent",
        "Language to which the word-form belongs to. Can be Catalan or Spanish",
        "Uniquely identifies the word-form in the questionnare",
        "Phonological transcription of the word-form in the IPA format",
        "Phonological transcription of the word-form in X-SAMPA format",
        "Syllables in the word-form",
        "Questionnaire sub-list in which the word-form appears",
        "Uniquely identifies every translation equivalent",
        "Levenshtein distance between the word-form and its translation",
        "Number of phonemes in the word-form",
        "Number of syllables in the word-form",
        "Lexical frequency in Zipf scores, from the English CHILDES corpora",
        "Sum of the frequency of the syllables in the word-form, expressed as counts per million tokens"
    ))
options(knitr.kable.NA = '--')

kable(items_dict,
      col.names = c(
          "Variable",
          "Type",
          "# missing",
          "# levels",
          "Mean",
          "Median",
          "SD",
          "Min",
          "Max",
          "Description"
      ))
```

## Participants

```{r participants-dict}
tar_load(participants)

participants_dict <- data_dict(participants) |> 
    mutate(comments = c(
        "Concept associated to the word-form, uniquely identifies every translation equivalent",
        "Language to which the word-form belongs to. Can be Catalan or Spanish",
        "Uniquely identifies the word-form in the questionnare",
        "Phonological transcription of the word-form in the IPA format",
        "Phonological transcription of the word-form in X-SAMPA format",
        "Syllables in the word-form",
        "Questionnaire sub-list in which the word-form appears",
        "Uniquely identifies every translation equivalent",
        "Levenshtein distance between the word-form and its translation",
        "Number of phonemes in the word-form",
        "Number of syllables in the word-form",
        "Lexical frequency in Zipf scores, from the English CHILDES corpora",
        "Sum of the frequency of the syllables in the word-form, expressed as counts per million tokens"
    ))
options(knitr.kable.NA = '--')

kable(items_dict,
      col.names = c(
          "Variable",
          "Type",
          "# missing",
          "# levels",
          "Mean",
          "Median",
          "SD",
          "Min",
          "Max",
          "Description"
      ))
:::

# Session info

```{r}
sessionInfo()
```