### 2023-07-26 _repro.qmd
Built site for gh-pages

---
---

We recommend two ways of reproducing or inspecting the code of this repository. One option is to clone the [Docker image](https://hub.docker.com/r/gongcastro/cognate-beginnings) we have pused to Dockerhub. A Docker image is just a virtual machine that contains all dependencies, code, and files necessary to run the code. Everything is already installed and setup, you will only have to open [https://localhost:8787](https://localhost:8787) in your browser, and a RStudio session will open for you. This option is more robust (dependencies are difficult to get right when using your own computer), but it also requires Docker to be installed.

A second option is the more traditional approach of downloading the repository (or cloning it, if you know your way through Git/GitHub), installing the necessary dependencies (we recommend using renv for R dependencies, other system dependencies are listed below), and running the code (we recommend using the targets workflow).