### 2023-07-27 index.qmd
Update website

---
title: "Cognate beginnings to bilingual lexical acquisition"
subtitle: "Lab notes"
short-title: "Cognate beginnings"
author:
  - name: Gonzalo Garcia-Castro
    corresponding: true
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    affiliations:
      - id: cbc
        name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias Fargas 25-27
        city: Barcelona
        region: Spain
        postal-code: "08005"
  - name: Daniela S. Avila-Varela
    orcid: 0000-0002-3518-8117
    affiliations: 
      - ref: cbc
      - id: ul
        name: Universidade de Lisboa
        department: Facultade de Letras
        address: Alameda da Universidade
        city: Lisboa
        region: Portugal
        postal-code: "1600-214"
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations: 
      - id: uam 
        name: Universidad Autónoma de Madrid
        department: Departamento de Psicología
        address: Iván Pavlov, 6
        city: Madrid
        region: Spain
        postal-code: "28049"
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations: 
      - ref: cbc
keywords:
  - lexical acquisition
  - vocabulary
  - bilingualism
  - item response theory
  - bayesian
link-citations: true
editor: source
warning: false
error: false
format:
  html:
    theme: _assets/style.scss
    page-layout: article
    code-fold: true
    smooth-scroll: true
toc: true
fig-dpi: 1000
sidebar: true
---

```{r setup}
#| label: setup
#| message: false
#| warning: false
#| echo: false
# load objects -----------------------------------------------------------------
targets::tar_config_set(
    store = here::here("_targets"),
    script = here::here("_targets.R")
)

yaml_metadata <- rmarkdown::yaml_front_matter(
    here::here("manuscript", "manuscript.qmd")
)

title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
thanks <- yaml_metadata$thanks

targets::tar_load(items)
targets::tar_load(participants)
targets::tar_load(responses)

library(dplyr)
library(tidyr)
library(knitr)

```

# Cognate beginnings to bilingual lexical acquisition

Gonzalo Garcia-Castro [:bird:](https://twitter.com/gongcastro) [:elephant:](https://fediscience.org/@gongcastro) • Universitat Pompeu Fabra

Daniela S. Avila Varela • Universidade de Lisboa

Ignacio Castillejo [:bird:](https://twitter.com/IndaloNacho) • Universidad Autónoma de Madrid

Nuria Sebastian-Galles • Universitat Pompeu Fabra

-------------------------------------------------

| Link                                                             | Contents                                                       |
|------------------------------------------------------------------|----------------------------------------------------------------|
| [Website](https://gongcastro.github.io/cognate.beginnings)       | Instructions for reproducibility, data dictionaries, lab notes |
| PsyArxiv                                                         | Preprint and figures                                           |
| [GitHub](https://github.com/gongcastro/cognate.beginnings)       | Code, preprint and figures                                     |
| [Open Science Framework](https://osf.io/hy984/)                  | Code, preprint, and results (model outputs)                    |
| [Docker](https://hub.docker.com/r/gongcastro/cognate-beginnings) | Docker image with reproducible RStudio session                 |

## Abstract

`r abstract`

## Acknowledgements

`r thanks`