### 2023-08-13 manuscript/appendix.qmd

---
csl: "_assets/apa7.csl"
bibliography: "_assets/references.bib"
link-citations: true
editor: source

fig-align: center
fig-dpi: 800
echo: false
warning: false
message: false

format:
  arxiv-pdf:
    linenumbers: true
    doublespacing: false
    runninghead: "Cognate beginnings to bilingual lexical acquistion (Appendix)"
    include-in-header: 
      text: |
        \usepackage{tipa}
        \usepackage{booktabs}
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggdist)
library(patchwork)

targets::tar_config_set(
    store = here::here("_targets"),
    script = here::here("_targets.R")
)

targets::tar_load_globals()

targets::tar_load(c(
    bvq_data, childes, items, model_summary,
    participants, responses
))

targets::tar_load(c(model_fit, model_convergence, model_ppcs))

targets::tar_load(c(posterior_doe_summary, posterior_doe_draws))

targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables, posterior_syllables_summary))

my_theme <- theme_minimal() +
    theme(
        panel.grid = element_line(
            colour = "grey",
            linetype = "dotted"
        ),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 12, colour = "black"),
        axis.text = element_text(colour = "black")
    )

theme_set(my_theme)
```

## Appendix A: model diagnostics {.appendix .unnumbered}

{{< include _appendix/_model-diagnostics.qmd >}}

{{< pagebreak >}}

## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_model-doe.qmd >}}

{{< pagebreak >}}

## Appendix C: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_syllable-frequency.qmd >}}

{{< pagebreak >}}

## References