### 2023-10-03 manuscript/appendix.qmd

---
csl: "_assets/apa7.csl"
bibliography: "_assets/references.bib"
link-citations: true
editor: source

fig-align: center
fig-dpi: 800
echo: false
warning: false
message: false

format:
  arxiv-pdf:
    linenumbers: true
    doublespacing: false
    runninghead: "Cognate beginnings to bilingual lexical acquistion (Appendix)"
    include-in-header: 
      text: |
        \usepackage{tipa}
        \usepackage{booktabs}
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggdist)
library(patchwork)

targets::tar_config_set(
    store = here::here("_targets"),
    script = here::here("_targets.R")
)

targets::tar_load_globals()

targets::tar_load(c(
    bvq_data, childes, 
    items,
    model_summary,
    participants, 
    responses
))

targets::tar_load(c(model_fit, 
                    model_convergence,
                    model_ppcs))

targets::tar_load(c(posterior_doe_summary, 
                    posterior_doe_draws))

targets::tar_load(c(syllables_data,
                    model_fit_syllables, 
                    post_syllables,
                    posterior_syllables_summary))

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(
        colour = "grey",
        linetype = "dotted"
    ),
    axis.line = element_line(colour = "black"),
    text = element_text(size = 12, colour = "black"),
    axis.text = element_text(colour = "black")
    )

theme_set(my_theme)

save_image <- function(file, ...) {
    suppressWarnings({
        invisible({
            devices <- c("png", "pdf", "svg", "tiff", "eps")
            paths <- file.path(paste(file, devices, sep = "."))
            purrr::map(paths,
                       \(x) ggsave(x, ...),
                       .progress = TRUE)
        })
    })
}

post_draws_list <- model_summary |>
    mutate(across(
        .value:.upper,
        \(x) ifelse(grepl("Intercept", .type),
                    plogis(x), x / 4
        )
    )) |>
    split(model_summary$.variable)
```

## Appendix A: model details {.appendix .unnumbered}

{{< include _appendix/_model-details.qmd >}}

{{< pagebreak >}}

## Appendix B: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_model-doe.qmd >}}

{{< pagebreak >}}

## Appendix C: frequency and language exposure as separate predictors {.appendix .unnumbered}

{{< include _appendix/_syllable-frequency.qmd >}}

{{< pagebreak >}}

## References